---
title: "Tidy Tuesday 2021"
subtitle: "Week 13"
output: github_document
author: "Lorena Abad"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
options(width=80)
knitr::opts_chunk$set(
  echo = TRUE,
  dev="CairoPNG", 
  warning = F, message = F, cache = F, 
  fig.showtext = T,
  tidy.opts=list(width.cutoff=80), 
  tidy=TRUE
)
```

# Tidy Tuesday week 13

## Libraries

```{r}
library(tidytuesdayR)
library(tidyverse)
library(sf)
library(patchwork)
library(rnaturalearth)
library(ragg)
```

## Get data

```{r}
tuesdata <- tt_load(2021, week = 13)
```

## Explore

```{r}
unvotes = tuesdata$unvotes
roll = tuesdata$roll_calls
issues = tuesdata$issues
```

### Filtering topic of interest

I found it interesting the outer space is discussed at the UN sessions, so I decided to explore the topics and how the votes were. I picked a single resolution after exploring.

```{r}
outer_space = roll %>% 
  filter(str_detect(tolower(descr), "no first placement")) %>% 
  left_join(unvotes) 

outer_clean = outer_space %>% 
  mutate(
    vote = factor(
      vote,
      levels = c("yes","abstain", "no")
    )
  )  %>% 
  mutate(
    topic = "No first placement of weapons in outer space",
    year_label = paste0("Year: ", format(date, "%Y")),
    session_label = paste0("Session: ", session)
  )
```

### Preparing spatial data

```{r}
no_countries = outer_clean %>% filter(vote == "no")
```

```{r}
# Inspiration from:
# https://stackoverflow.com/questions/43207947/whole-earth-polygon-for-world-map-in-ggplot2-and-sf
# CRS is the one used for the official UN logo
crs = "+proj=aeqd +lat_0=85 +lon_0=0 +x_0=0 +y_0=0 +ellps=GRS80 +units=m +no_defs"
# Get countries from rnaturalearth
world = ne_countries(scale = "medium", returnclass = "sf") %>% 
  select(admin, iso_a2)
# Draw a sphere to plot countries on top
# It will represent the oceans
sphere = st_graticule(ndiscr = 10000, margin = 10e-6) %>%
  st_transform(crs = crs) %>%
  st_convex_hull() %>%
  summarise(geometry = st_union(geometry))

countries_geometry = outer_clean %>% 
  left_join(world, by = c("country_code" = "iso_a2")) %>% 
  st_as_sf()
```


## Plot

### Setting themes

```{r}
# Load fonts
extrafont::loadfonts('win')
theme_set(theme_minimal(base_family = "Futura LT"))
theme_update(
  axis.title = element_blank(),
  axis.ticks = element_blank(),
  axis.text = element_blank(),
  panel.grid = element_blank()
)
```

### Setting colors

```{r}
votepal = c(
  "yes" = "deepskyblue3",
  "abstain" = "khaki", 
  "no" = "firebrick1"
)
```

### Caption text

```{r}
caption_text = paste0(
  "Data: ",
  "Harvard's Dataverse. Voeten, 2013. ",
  "Visualization: ", 
  "@loreabad6"
  )
```

### Base plot

```{r, fig.height = 35, fig.width = 10}
g = ggplot(
    no_countries,
    aes(y = vote, x = country_code, label = country_code)
  ) +
  # with_outer_glow(
    geom_point(color = votepal['no'], shape = 20, size = 2.5 #),
    # color = "white", sigma = 10 #, ignore_background = F
  ) +
  geom_text(
    color = "white",
    size = 2.5,
    nudge_x = 0, nudge_y = 0.4,
    family = "Futura LT"
  ) +
  scale_y_discrete(
    drop = F,
    expand = expansion(mult = c(0,0), add = c(0,0))
  ) +
  facet_wrap(
    ~ session_label + year_label,
    nrow = 1,
    labeller = label_wrap_gen(width = 25, multi_line = TRUE)
  ) +
  coord_polar(clip = "off") +
  theme(
    text = element_text(color = "white"),
    legend.position = "none",
    panel.spacing = unit(0, "cm"),
    plot.margin = margin(rep(5, 4)),
    plot.background = element_rect(
      color = "black",fill = "black"
    ),
    panel.background = element_rect(
      color = "black", fill = "black"
    ),
    strip.text = element_text(color = "white"),
    strip.text.y.left = element_text(hjust = 1, angle = 0),
    strip.background = element_blank(),
    plot.title.position = "plot",
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, size = 9),
    plot.caption.position = "plot",
    plot.caption = element_text(hjust = 0, size = 8)
  ) +
  labs(
    title = "UN members and the Prevention of an Arms Race in Outer Space\nResolution: No first placement of weapons in outer space",
    subtitle = str_wrap("The number of UN member countries voting against policy preventing the first placement of weapons in outer space has increased almost four times since the resolution was introduced in 2014. Countries voting against the policy are represented as points orbiting the earth.", 100),
    caption = caption_text
  )
```

### Legend plot

```{r}
legend = no_countries %>% 
  select(country, country_code, vote) %>% 
  distinct() %>% 
  arrange(country_code) %>% 
  mutate(
    label = paste0(
      # country_code, "\n",
      str_replace_all(country, " ", "\n")
    ), 
    id = row_number()
) 

# From https://www.r-graph-gallery.com/296-add-labels-to-circular-barplot.html
# ----- This section prepare a dataframe for labels ---- #
# Get the name and the y position of each label

# calculate the ANGLE of the labels
number_of_points = nrow(legend)
angle =  90 - 360 * (legend$id-0.5) / number_of_points
# I substract 0.5 because the letter must have the angle of the center of the bars. Not extreme right(1) or extreme left (0)
 
# calculate the alignment of labels: right or left
# If I am on the left part of the plot, my labels have currently an angle < -90
# flip angle BY to make them readable
legend2 = legend %>% 
  mutate(
    hjust = ifelse(angle < -90, 1, 0),
    angle = ifelse(angle < -90, angle + 180, angle)
  )
 # ----- ------------------------------------------- ---- #
 

l = ggplot(legend2, aes(y = vote, x = country_code)) +
  geom_point(color = "white", shape = 20, size = 1.5
  ) +
  geom_text(
    aes(hjust = hjust, label = label),
    color = "white",
    size = 1.5,
    nudge_x = 0, nudge_y = 0.6,
    family = "Futura LT",
    angle = legend2$angle
  ) +
  scale_y_discrete(
    drop = F,
    expand = expansion(mult = c(0,0), add = c(0,0))
  ) + 
  coord_polar(clip = "off") 
```

### World plot

```{r}
w = ggplot() +
  geom_sf(
    data = sphere,
    fill = "gray10", color = NA, alpha = 0.8
  ) +
  geom_sf(
    data = filter(world, admin != "Antarctica"),
    fill = "grey50", size = 0.001, color = NA
  ) +
  geom_sf(data = countries_geometry, 
          aes(fill = vote), color = NA
  ) +
  coord_sf(crs = crs, clip = "off") +
  scale_fill_manual(
    "Vote:",
    values = votepal,
    na.value = "grey50"
  ) +
  theme(
    text = element_text(color = "white"),
    plot.background = element_rect(fill = "transparent",
                                   color = "transparent"), 
    panel.background = element_rect(fill = "transparent",
                                    color = "transparent"),
    plot.margin = margin(rep(0.5, 4)),
    panel.spacing = unit(1.5, "cm"),
    legend.position = "top",
    strip.background = element_rect(fill = "transparent",
                                    color = NA),
    strip.text = element_blank(),
    legend.background = element_blank()
  ) +
  facet_wrap(~ session_label, nrow = 1) 
```

### Arrange patchwork

```{r}
layout = c(
  area(1,1,20,60),
  area(1,3,20,58)
)

g_inset = g + 
  inset_element(
    l,
    left = 0.8, bottom = 1.3,
    right = 1, top = 1.8
  )

p = g_inset + w + plot_layout(design = layout)
```

### Save plot

```{r}
agg_png(
  filename = "plot/2021_week_13.png",
  width = 24, height = 11,
  units = "cm", res = 300
)
p
invisible(dev.off())
knitr::plot_crop("plot/2021_week_13.png")
```
