---
title: "Tidy Tuesday 2020"
subtitle: "Week 37"
output: github_document
author: "Lorena Abad"
editor_options: 
  chunk_output_type: console
---

```{r setup, echo = F}
knitr::opts_chunk$set(dev="CairoPNG", warning = F, message = F)
```

# Tidy Tuesday week 37

This is my very first submission to Tidy Tuesday! Pretty excited I got to explore a sitcom I basically grew up with!

## Get data

```{r}
library(tidytuesdayR)
tuesdata <- tidytuesdayR::tt_load(2020, week = 37)
```

## Explore

```{r}
library(dplyr)
tuesdata$friends %>% glimpse()
tuesdata$friends %>% summary()
tuesdata$friends_info %>% glimpse()
tuesdata$friends_emotions %>% glimpse()
```

After exploring I decided to keep it simple, and will look at the number of lines by main character. 

## Extract main speakers 

Naturally, they should be the main 6 characters, other than "Scene Directions" which I filtered out.

```{r}
characters = tuesdata$friends %>% 
  filter(speaker != 'Scene Directions') %>% 
  group_by(speaker) %>% 
  summarize(no_of_lines = n()) %>% 
  top_n(6, no_of_lines)
```

## Get some extra data
I was very amused by the [ceros.com](https://www.ceros.com/originals/friends-scripts-25th-anniversary-catchphrase-scenes-quotes/) analysis on the Tidy Tuesday repo and thought would pick a particular analysis they already did: "Number of lines between characters". 

```{r}
library(tibble)
lines_between_characters = tribble(
  ~to, ~from, ~no_of_lines,
  'Rachel Green', 'Ross Geller', 3796,
  'Monica Geller', 'Chandler Bing', 3201,
  'Chandler Bing', 'Joey Tribbiani', 2563,
  'Rachel Green', 'Phoebe Buffay', 1776,
  'Rachel Green', 'Monica Geller', 1756,
  'Monica Geller', 'Phoebe Buffay', 1684,
  'Rachel Green', 'Joey Tribbiani', 1662,
  'Joey Tribbiani', 'Ross Geller', 1589,
  'Chandler Bing', 'Ross Geller', 1295,
  'Joey Tribbiani', 'Phoebe Buffay', 1204,
  'Phoebe Buffay', 'Ross Geller', 1068,
  'Monica Geller', 'Ross Geller', 921,
  'Monica Geller', 'Joey Tribbiani', 920,
  'Chandler Bing', 'Phoebe Buffay', 795,
  'Chandler Bing', 'Rachel Green', 670
)
```

With the characters and the lines between them, I decided to do a network representation of the data with `tidygraph` and `ggraph`.

## Setting up the graph

### Preparing the "nodes"
First, some data wrangling. Split first and last name, place the characters in an order I picked (really trial and error from the final plot), and add some extra info for the plotting. 

```{r, fig.showtext = TRUE}
characters = characters %>% 
  tidyr::separate(speaker, c('first_name', 'last_name'), remove = F) %>% 
  mutate(order = c(2,4,5,3,1,6)) %>% 
  arrange(order) %>% 
  mutate(speaker = factor(speaker, ordered = T)) %>%   
  mutate(letter = unlist(strsplit('FRIEND',"")))
```

### Building the graph

```{r}
library(ggraph)
library(tidygraph)

friends_graph = tbl_graph(
  nodes = characters, edges = lines_between_characters, 
  directed = T, node_key = 'speaker')
```

## And plotting!
```{r, fig.height = 5, fig.width = 15}
# Fonts sources:
## Main sitcom font: https://www.ffonts.net/Friends.font.download#
## Credits font: https://fontsgeek.com/fonts/Lydian-Regular

# Define colors
colors = c('#FF4238', '#42A2D6', '#FFDC00', '#FF4238', '#FFDC00', '#42A2D6')

# Load fonts
extrafont::loadfonts('win')

# Plot
g = ggraph(friends_graph, layout = 'linear') +
  geom_edge_arc(
    aes(color = no_of_lines, width = no_of_lines, alpha = stat(index)),
    strength = 0.3,
  ) +
  scale_edge_alpha(guide = 'none') +
  scale_edge_color_gradient("Lines between characters", low = 'white', high = '#f6c719', guide = 'legend') +
  scale_edge_width("Lines between characters", range = c(0.3, 1.5)) +
  geom_node_point(aes(color = speaker, size = no_of_lines), shape = 19, stroke = 2, show.legend = T) +
  geom_node_text(
    aes(label = letter), nudge_x = -0.5, size = 7, 
    color = 'white', family = 'Friends', show.legend = F
  ) +
  geom_node_text(
    aes(label = first_name), size = 4, fontface = 'bold',
    color = '#1c0e01', family = 'Lydian', show.legend = F
  ) +
  annotate('text', x = 6.5, y = 0, label = 'S', color = 'white', size = 7, family = 'Friends') +
  scale_size("Lines by character", range = c(18, 35)) +
  scale_color_manual(breaks = characters$speaker, values = colors, guide = 'none') +
  scale_y_discrete(expand = expansion(0.1, 0)) +
  scale_x_discrete(expand = expansion(0.1, 0)) +
  guides(
    edge_color = guide_legend(override.aes = list(shape = NA)),
    width = guide_legend(override.aes = list(shape = NA)),
    size = guide_legend(order = 1, 
                        override.aes = list(color = 'white', size = c(2,8,15)))
    ) +
  labs(
    # title = "The one with all the lines",
    caption = "Data: {friends} R package and ceros.com. Created by: @loreabad6"
  ) +
  theme(
    text = element_text(family = 'Friends', color = 'white', size = 9),
    plot.caption = element_text(family = 'Lydian', size = 8, angle = 0, hjust = 0, vjust = 0.5),
    plot.caption.position = 'plot',
    legend.position = 'bottom',
    legend.background = element_rect(fill = 'transparent'),
    legend.text = element_text(size = 6),
    legend.key = element_rect(fill = 'transparent'),
    legend.box = 'vertical', legend.direction = 'horizontal',
    legend.spacing = grid::unit(2, 'mm'),
    panel.background = element_rect(fill = '#1c0e01'),
    plot.background = element_rect(fill = '#1c0e01')
  )
g
```


## Save plot
```{r}
ggsave(g, filename = "plot/2020_week_37.png", device = "png", width = 30, height = 15, units = "cm")
```

